---
title: "ClimDist_TigerMtn"
author: "Elizabeth R. Milano"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
```

## Climate Distance mapper - Tiger Mountain example

**Purpose:** Climate Distance algorithm for current and future analogs.
**Background:** Climate Distance project aims to identify contemporary
climate analog locations for future climate scenarios at focal points of
interest based on derived climate variables, plant survey data, and biotic community classifications (hereafter biomes). This
document identifies analogs for the Tiger Mountain target location using the Forest Inventory and Analysis dataset, and Brown's biomes.  

**Data:**   

* Background point dataset: This is used to define analog thresholds and
determine `mean` and `sd` of derived climate variables for standardization
purposes. The dataset is comprised of 988 lat/long locations, each point
was then sampled along an artificial elevation cline to create a set of
50,388 observations. This analysis was done in a previous script and
does not change across reference databases.   
* Reference point dataset: For each inventory database (FIA, biome,
ect.), contemporary climate data was pulled from
`https://charcoal2.cnre.vt.edu//climate/customData/` for all lat/long
occurrences. 
Biome classification can be found at
`http://www.peter.unmack.net/biotic/`.
FIA data https://apps.fs.usda.gov/fia/datamart/datamart.html.
Data accessed Nov. 2022.   
* Target dataset: Contains 13 future climate projections and contemporary data from source
above for Tiger Mountain. *Note: This script is set up to loop through a list of target locations and create separate output files for each, however we will only focus on a single location for sake of example*   

**Analysis:** We are interested in comparing plant species abundance and
composition between contemporary climate analogs and future climate
analogs. For both contemporary and future scenarios the nearest neighbor
imputation with `yaimpute` function is run twice. First on the set of
reference points to identify top 10,000 nearest neighbor distances which
are carried forward to a second imputation with the added target
contemporary or future occurrences. Analogs are identified under `Edist <0.84` threshold.

**Results files:** Text files contain list of analogs for each location and
climate scenario, data for each analog includes lat/long, Edist (climate
distance), km (geographic distance from focal location), and either
species occurrence information or biome classification. Summary tables follow results output. 
Species comparisons and maps are generated in separate script.

```{r env}
setwd("/Users/elizabethmilano/Library/CloudStorage/Box-Box/01. elizabeth.milano Workspace/ClimDist/github/scripts")
rm(list=ls())

library(yaImpute) # nearest neighbor
library(gower)
library(geosphere) # calculate km distance between geo points
library(tidyverse) # tidy data
```

## Load data

### Background data stats

Used for standardization and thresholds.\

```{r stats_data}
stats <- read.csv("../data/ClimDist_stats.csv")
```

### Reference data

#### Current climate

Reference dataset is a combined set of points from FIA and biome data. A
new PointID index was generated for the biome and FIA points by starting the count at
30,000 and 500,000 to avoid duplication with the AIM dataset (not used for Tiger Mountain).

```{r ref_data}
# Current climate for reference points
#aim.data <- read.table("../data/AIM_currentClim_derived.txt",header=T,sep="",na.strings=".") %>%
#  rename(mindd0 = mmindd0) %>%
#  select(-smrsprpb, -smrpb) %>%
#  mutate(zone = NA,
#         PointID_biome = NA,
#         PointID_fia = NA)

biome.tmp <- read.table("../data/BRupdate.txt",header=T,sep="",na.strings=".") %>%
  rename(PointID_biome = PointID,
         smrp = sump) 
n_row.b <- nrow(biome.tmp)
biome.data <- biome.tmp %>%
  mutate(PointID = seq(30001,(30000+n_row.b)), #index accounts for AIM dataset, if using.
         PointID_fia = NA)


fia.tmp1 <- read.table("../data/FIA_08.24.15.txt",header=T,sep="",na.strings=".") %>%
  rename(PointID_fia = newid,
         smrp = sump)
n_row.f <- nrow(fia.tmp1)
fia.tmp2 <- fia.tmp1 %>%
  mutate(PointID = seq(500001,(500000+n_row.f)), #index accounts for AIM and biome datasets, if using.
         PointID_biome = NA,
         zone = NA)
fia.data <- fia.tmp2 %>%
  select(PointID, long, lat, elev, mat, map, gsp, dd5, dd0, gsdd5, mtcm, mmin, mtwm, mmax, sday, fday, ffp, d100, mindd0, winp, smrp, sprp,
         PointID_biome, zone, PointID_fia)

ref.data <- rbind(#aim.data, 
                  biome.data, 
                  fia.data) %>%
  select(-PointID_biome, -PointID_fia)

```

#### Vegetation and biomes

```{r veg_data}
# Vegetation
#aim.species <- read.table("../data/BLMfull_withnox.csv",header=T,sep=",",na.strings="NA") 

#biome
biome <- biome.data %>%
  select(PointID, PointID_biome, zone)
biome.code <- read.delim("../data/Brown_biome_code.txt")

#FIA
fia.species <- fia.tmp2 %>%
  select(-mat, -map, -gsp, -dd5, -dd0, -gsdd5, -mtcm, -mmin, -mtwm, -mmax, -sday, -fday, -ffp, -d100, -mindd0, -winp, -smrp, -sprp,
         -PointID_biome, -zone, -long, -lat, -elev, -PIGL.1)
# FIA species
fia.species.lookup <- read.csv("../data/FIA_master_list.csv", na.strings = "NA") %>%
  unite("latin", Genus, Species, Subspecies, Variety, na.rm = TRUE, remove = FALSE) %>%
  rename(species_code = PLANTS_Code)
```

### Target data

#### Future climate

Pull `*_y2060_derived.txt` files from
`https://charcoal2.cnre.vt.edu//climate/customData/`.

```{r future_data}

# Target points
target.future.files <- list.files(path="../data/charcoal",full.names=TRUE, pattern = "y2060_derived.txt$")

# list futures
future.table <- data.frame(Filepath = target.future.files) %>%
  separate(col = Filepath, sep = "/", into = c(NA, NA, NA, "File")) %>%
  separate(col = File, sep = "_", into = c("Model", "Scenario", "Year"), remove = FALSE)

target.future.data <- read_delim(target.future.files, col_names = TRUE, delim=" ", na=".", show_col_types = FALSE) %>% 
  bind_rows() %>%
  group_by(PointID) %>%
  summarise_all(mean)%>%
  rename(mindd0 = mmindd0)
write.csv(target.future.data, "../output/future/future_climate_target.csv", row.names = FALSE)


# list focal points
fp.table <- target.future.data %>%
  select(PointID, long, lat, elev) %>%
  rename("Longitude W" = "long",
         "Latitude N" = "lat",
         "Elevation m" = "elev")

```

#### Contemporary climate

```{r current_data}

target.current.data <- read.table("../data/charcoal/currentClim_derived.txt",header=T,sep="",na.strings=".") %>%

  rename(mindd0 = mmindd0)
write.csv(target.current.data, "../output/current/current_climate_target.csv", row.names = FALSE)


```

#### Target data summary

```{r table1, echo=FALSE}
knitr::kable(future.table, format = "simple", caption = "Table 1. Future climate scenarios.")
```

```{r table2, echo=FALSE}
knitr::kable(fp.table, format = "simple", caption = "Table 2. Target point location(s).")
```

## Functions

1.  Create derived climate variables

Climate variable definitions:   
All temperature is in degrees C and precipitation in mm.    

Raw variables:   
**dd5** = Degree-days \>5 (based on mean monthly temperature).  
**mat** = Mean annual temperature.   
**d100** = Julian date the sum of degree-days >5 reaches 100.  
**map** = Mean annual precipitation.  
**gsp** = Growing season precipitation, April to September.  
**smrp** = Summer precipitation: jul+aug.
**sprp** = Spring precipitation: 
**winp** = Winter precipitation: nov+dec+jan+feb.   
**mmax** = Mean maximum temperature in the warmest month.   
**mtcm** = Mean temperature in the coldest month.   
The following are not included directly in climate distance, but used to calculate
derived variables:   
*mmindd0* = Degree-days \<0 (based on mean minimum monthly temperature); a.k.a. *mindd0*.   
*gsdd5* = Degree-days \>5 accumulating within the frost-free period.   
*mtwm* = Mean temperature in the warmest month.    

Derived variables:  
**tdiff** = Mean temp warmest month minus coldest month, `mtwm - mtcm`.   
**pratio** = Ratio of summer precipitation to total precipitation, `gsp / map`.   
**adi** = Annual dryness index, `sqrt(dd5)/map`.   
**sdi** = Summer dryness index, `sqrt(gsdd5)/gsp`.  
**adimindd0** = `adi * (mindd0 + 1)`   
**sdimindd0** = `(sqrt(gsdd5) / gsp) * (mindd0 + 1))`   
**mapdd5** = `(map * dd5) / 1000`  
**gspdd5** = `(gsp * dd5) / 1000`  
**smratio** = `smrp/gsp` 
```{r functions1}
der.vars <- function(df) {
  output <- df %>%
    mutate(tdiff = mtwm - mtcm,
           pratio = gsp / map,
           adi = (dd5 ** 0.5) / map,
           sdi = (gsdd5 ** 0.5) / gsp,
           mapdd5 = (map * dd5) / 1000,
           gspdd5 = (gsp * dd5) / 1000,
           adimindd0 = adi * (mindd0 + 1),
           sdimindd0 = ((gsdd5 ** 0.5) / gsp) * (mindd0 + 1),
           smratio = smrp/gsp) 
  return(output)
}
```



2.  Standardize new variables with `stats` dataset




```{r functions2}
# Standardize vars with "stats"
std.stats <- function(df){
  output <- df %>%
    dplyr::mutate(dd5 = (dd5 - stats$Xdd5) / stats$Sdd5,
           mat = (mat - stats$Xmat) / stats$Smat,
           d100 = (d100 - stats$Xd100) / stats$Sd100,
           map = (map - stats$Xmap) / stats$Smap,
           gsp = (gsp - stats$Xgsp) / stats$Sgsp,
           winp = (winp - stats$Xwinp) / stats$Swinp,
           mmax = (mmax - stats$Xmmax) / stats$Smmax,
           tdiff = (tdiff - stats$Xtdiff) / stats$Stdiff,
           pratio = (pratio - stats$Xpratio) / stats$Spratio,
           adi = (adi - stats$Xadi) / stats$Sadi,
           sdi = (sdi - stats$Xsdi) / stats$Ssdi,
           adimindd0 = (adimindd0 - stats$Xadimindd0) / stats$Sadimindd0,
           sdimindd0 = (sdimindd0 - stats$Xsdimindd0) / stats$Ssdimindd0,
           mtcm = (mtcm - stats$Xmtcm) / stats$Smtcm,
           mapdd5 = (mapdd5 - stats$Xmapdd5) / stats$Smapdd5,
           gspdd5 = (gspdd5 - stats$Xgspdd5) / stats$Sgspdd5,
           smrp = (smrp - stats$Xsmrp)/stats$Ssmrp, 
           sprp = (sprp -stats$Xsprp)/stats$Ssprp, 
           smratio = (smratio - stats$Xsmratio)/stats$Ssmratio 
          ) %>%
    select(PointID, long, lat, elev, dd5, mat, d100, map, gsp, winp, mmax, tdiff, pratio, adi, sdi, adimindd0, sdimindd0, mtcm, mapdd5, gspdd5, smrp, sprp, smratio)
  return(output)
}
```

## Data Transformation

Apply functions above to prepare files.

```{r transform}
# reference data
ref.vars <- der.vars(ref.data)
ref <- std.stats(ref.vars)

# target future data
target.future.vars <- der.vars(target.future.data)
target.future <- std.stats(target.future.vars)

# target current data
target.current.vars <- der.vars(target.current.data)
target.current <- std.stats(target.current.vars)
```

## Find future climate analogs

-   Use `yaImpute` to calculate nearest neighbors for current climate
    reference points.\
-   Then apply to future climate of focal points.\
-   Select analog points with `Edist < 0.9` (from threshold
    calculation).\
-   Calculate distance in km from focal point to analog point and
    classify as strong, moderate, or weak.\
-   Output `.txt` files for each focal location with analog points and
    current species occurrences or biome classifications.\
      (+ AIM `/output/future/future_analog_[location]_AIM.txt`) *not used for Tiger Mtn*\
       + Biome `/output/future/future_analog_[location]_biome.txt`\
       + FIA `/output/future/future_analog_[location]_FIA.txt`\

```{r yaimpute_f}
#aim.future <- list()
biome.future <- list()
fia.future <- list()
#clim.future <- list()

 for (irow in 1:nrow(target.future))
 {

 trg <- target.future[irow,]
 
 # set focal point lat/long
 trg.long <- trg$long
 trg.lat <- trg$lat
 trg.elev <- trg$elev
 
 # yai format needs rownames
 rownames(ref) = paste("R",sprintf(fmt="%4.4d",1:nrow(ref)),sep="")
 rownames(trg) = paste("",sprintf(fmt="%4.4d",1:nrow(trg)),sep="")
 
 # yaImpute
 yai = yai (x=ref[,5:ncol(ref)],method="raw",noRefs=TRUE,k=2500)
 yai2 = newtargets(yai,newdata=trg)
 
 # pull results
 ids = yai2$neiIdsTrg[,1:2500] # identifies R_number from ref
 dsts = round(yai2$neiDstTrg[,1:2500],2)

  # format output
#	aim.future[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
#	  filter(Edist <= 1.75) %>%
#	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
#	  select(PointID, long, lat, elev, Edist) %>%
#	  mutate(Edist_thresh = case_when(
#      Edist <= 0.6 ~ "strong",
#      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
#      Edist > 0.72 & Edist <= 0.84 ~ "weak",
#      Edist > 0.84 ~ "under_thresh",
#      TRUE ~ "NA")) %>%
#	  rowwise() %>% # needed for km calculation
#	  mutate(elev_diff = trg.elev - elev,
#	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
#	  left_join(aim.species) %>%
#	  drop_na() %>% 
#	  select_if(~ !is.numeric(.) || sum(.) != 0) # remove species columns without info
#
#    write.table (aim.future[[trg$PointID]], paste0("../output/future/","future_analog_", trg$PointID, "_AIM.txt"),row.names=FALSE)

    biome.future[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
	  filter(Edist <= 1.75) %>%
	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
	  select(PointID, long, lat, elev, Edist) %>%
	  mutate(Edist_thresh = case_when(
      Edist <= 0.6 ~ "strong",
      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
      Edist > 0.72 & Edist <= 0.84 ~ "weak",
      Edist > 0.84 ~ "under_thresh",
	    TRUE ~ "NA")) %>%
	  rowwise() %>% # needed for km calculation
	  mutate(elev_diff = trg.elev - elev,
	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
	  left_join(biome) %>%
    left_join(biome.code) %>%
    drop_na()

    write.table (biome.future[[trg$PointID]], paste0("../output/future/","future_analog_", trg$PointID, "_biome.txt"),row.names=FALSE)
    
    fia.future[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
	  filter(Edist <= 1.75) %>%
	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
	  select(PointID, long, lat, elev, Edist) %>%
	  mutate(Edist_thresh = case_when(
      Edist <= 0.6 ~ "strong",
      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
      Edist > 0.72 & Edist <= 0.84 ~ "weak",
      Edist > 0.84 ~ "under_thresh",
      TRUE ~ "NA")) %>%
	  rowwise() %>% # needed for km calculation
	  mutate(elev_diff = trg.elev - elev,
	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
	  left_join(fia.species) %>%
	  drop_na() %>% 
	  select_if(~ !is.numeric(.) || sum(.) != 0) # remove species columns without info

    write.table (fia.future[[trg$PointID]], paste0("../output/future/","future_analog_", trg$PointID, "_FIA.txt"),row.names=FALSE)
    
#    ## output climate vars for all analogs
#    clim.future[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
#	  filter(Edist <= 1.75) %>%
#	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
##	  select(PointID, long, lat, elev, Edist) %>%
#	  mutate(Edist_thresh = case_when(
#      Edist <= 0.6 ~ "strong",
#      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
#      Edist > 0.72 & Edist <= 0.84 ~ "weak",
#      Edist > 0.84 ~ "under_thresh",
#	           TRUE ~ "NA"),
#	   analog_dataset = case_when(
#	    PointID <= 30000 ~ "AIM",
#	    PointID > 30000 & PointID <= 500000 ~ "Biome",
#	    PointID > 500000 ~ "FIA",
#	    TRUE ~ "NA")) %>%
#	  rowwise() %>% # needed for km calculation
#	  mutate(elev_diff = trg.elev - elev,
#	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
#	  drop_na()
#
#    write.table (clim.future[[trg$PointID]], paste0("../output/future/","future_analog_", trg$PointID, "_climate.txt"),row.names=FALSE)
}

#clim.future.df <-  bind_rows(clim.future, .id = "PlotID") %>%
#  filter(Edist < 0.9)
#write.csv(clim.future.df, "../output/future/future_GB_trans_climate.csv", row.names = FALSE)
```

```{r table3, echo=FALSE}
#knitr::kable(aim.future[[8]][1:5,1:15], format = "simple", caption = "Table 3. Example format for output file for future climate analogs with AIM species composition.")
```


```{r table4, echo=FALSE}
knitr::kable(biome.future[[1]][1:6,], format = "simple", caption = "Table 4. Example format for output file for future climate analogs with biome classification.")
```


```{r table5, echo=FALSE}
knitr::kable(fia.future[[1]][1:6,], format = "simple", caption = "Table 5. Example format for output file for future climate analogs with FIA species composition.")
```

## Find current climate analogs

-   Same process as future climate analogs except using the contemporary
    climate data as target dataset.\
-   Output files:\
      (+ AIM `/output/current/current_analog_[location]_AIM.txt`) *not used for Tiger Mtn*\
       + Biome `/output/current/current_analog_[location]_biome.txt`\
       + FIA `/output/current/current_analog_[location]_FIA.txt`\

```{r yaimpute_c}
#aim.current <- list()
biome.current <- list()
fia.current <- list()

 for (irow in 1:nrow(target.current))
 {

 trg <- target.current[irow,]

 # set focal point lat/long
 trg.long <- trg$long
 trg.lat <- trg$lat
 trg.elev <- trg$elev

 # yai format needs rownames
 rownames(ref) = paste("R",sprintf(fmt="%4.4d",1:nrow(ref)),sep="")
 rownames(trg) = paste("",sprintf(fmt="%4.4d",1:nrow(trg)),sep="")
 
 # yaImpute
 yai = yai (x=ref[,5:ncol(ref)],method="raw",noRefs=TRUE,k=2500)
 yai2 = newtargets(yai,newdata=trg)
 
 # pull results
 ids = yai2$neiIdsTrg[,1:2500] # identifies R_number from ref
 dsts = round(yai2$neiDstTrg[,1:2500],2)

  # format output
#	aim.current[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
#	  filter(Edist <= 1.75) %>%
#	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
#	  select(PointID, long, lat, elev, Edist) %>%
#	  mutate(Edist_thresh = case_when(
#      Edist <= 0.6 ~ "strong",
#      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
#      Edist > 0.72 & Edist <= 0.84 ~ "weak",
#      Edist > 0.84 ~ "under_thresh",
#	    TRUE ~ "NA")) %>%
#	  rowwise() %>% # needed for km calculation
#	  mutate(elev_diff = trg.elev - elev,
#	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
#	  left_join(aim.species) %>%
#	  drop_na() %>% 
#	  select_if(~ !is.numeric(.) || sum(.) != 0) # remove species columns without info
# 
#    write.table (aim.current[[trg$PointID]], paste0("../output/current/","current_analog_", trg$PointID, "_AIM.txt"),row.names=FALSE)

    
    biome.current[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
	  filter(Edist <= 1.75) %>%
	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
	  select(PointID, long, lat, elev, Edist) %>%
	  mutate(Edist_thresh = case_when(
      Edist <= 0.6 ~ "strong",
      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
      Edist > 0.72 & Edist <= 0.84 ~ "weak",
      Edist > 0.84 ~ "under_thresh",
	    TRUE ~ "NA")) %>%
	  rowwise() %>% # needed for km calculation
	  mutate(elev_diff = trg.elev - elev,
	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
	  left_join(biome) %>%
    left_join(biome.code) %>%
    drop_na()

    write.table (biome.current[[trg$PointID]], paste0("../output/current/","current_analog_", trg$PointID, "_biome.txt"),row.names=FALSE)
    
    fia.current[[trg$PointID]] <- data.frame(Edist=dsts, id=ids) %>%
	  filter(Edist <= 1.75) %>%
	  left_join(ref %>% rownames_to_column("rowname"), by=c("id"="rowname")) %>% # join by 'R' rowname of ref
	  select(PointID, long, lat, elev, Edist) %>%
	  mutate(Edist_thresh = case_when(
      Edist <= 0.6 ~ "strong",
      Edist > 0.6 & Edist <= 0.72 ~ "moderate",
      Edist > 0.72 & Edist <= 0.84 ~ "weak",
      Edist > 0.84 ~ "under_thresh",
    TRUE ~ "NA")) %>%
	  rowwise() %>% # needed for km calculation
	  mutate(elev_diff = trg.elev - elev,
	         km = (distm(c(trg.long, trg.lat), c(long,lat), fun = distGeo)/1000)[1]) %>% # km distance between focal and analog lat/long
	  left_join(fia.species) %>%
	  drop_na() %>% 
	  select_if(~ !is.numeric(.) || sum(.) != 0) # remove species columns without info

    write.table (fia.current[[trg$PointID]], paste0("../output/current/","current_analog_", trg$PointID, "_FIA.txt"),row.names=FALSE)
        
}

```

## Analog summaries
*Code can be used for multiple target locations, here we will only look at Tiger Mountain FIA and biome results*. 

### Biome dataset

Load results files.  

```{r biome_results}
# Biome
current.b.files <- list.files(path="../output/current",full.names=TRUE, pattern = "_biome.txt")
current.b <- list()
for (file in current.b.files){
  name.tmp <- gsub(pattern = "../output/current/current_analog_", replacement = "", x = file)
  name <- gsub(pattern = "_biome.txt", replacement = "", x = name.tmp)
  current.b[[name]] <- read.delim(file = file, header=T, sep="", na.strings="NA") %>%
    mutate(scenario = "current",
           location = name)
}

future.b.files <- list.files(path="../output/future",full.names=TRUE, pattern = "_biome.txt")
future.b <- list()
for (file in future.b.files){
  name.tmp <- gsub(pattern = "../output/future/future_analog_", replacement = "", x = file)
  name <- gsub(pattern = "_biome.txt", replacement = "", x = name.tmp)
  future.b[[name]] <- read.delim(file = file, header=T, sep="", na.strings="NA") %>%
    mutate(scenario = "future",
           location = name) 
}
```

Count analogs.  

```{r biome2}
## Current analogs
# remove empty elements from list
current.b <- current.b[map(current.b, ~dim(.)[1]) > 0]
# bind rows from list to dataframe    
current.b.df <-  bind_rows(current.b, .id = "PlotID") %>%
  filter(Edist <= 0.84) 
  
## Future analogs
# remove empty elements from list
future.b <- future.b[map(future.b, ~dim(.)[1]) > 0]
# bind rows from list to dataframe    
future.b.df <-  bind_rows(future.b, .id = "PlotID") %>%
  filter(Edist <= 0.84) 

b.plot <- current.b.df %>%
  bind_rows(future.b.df) %>%
  select(PlotID, scenario, Edist, Edist_thresh, elev, elev_diff, km)

b.bar <- ggplot(b.plot, aes(x=scenario, fill=Edist_thresh)) +
  geom_bar(position="dodge") +
  labs(title="Biome analog count") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) 
b.bar

```

Compare current and future.

```{r biome3}
biome.sup.plots <- current.b.df %>%
  bind_rows(future.b.df) %>%
 # filter(PlotID %in% plots) %>%
  select(-PointID, -long, -lat, -Edist, -Edist_thresh, -elev, -elev_diff, -km, -PointID_biome) %>%
  group_by(scenario, PlotID, biome_description) %>% # if plots > 1, group by PlotID
  summarize(n=n()) %>%
  mutate(freq = n/sum(n)) %>%
  pivot_wider(id_cols =  "biome_description", names_from = c("PlotID","scenario"), values_from = "freq")
biome.sup.plots


biome.sup.plots.stats <- current.b.df %>%
  bind_rows(future.b.df) %>%
#  filter(PlotID %in% plots) %>%
  select(scenario, elev, elev_diff, km, PlotID) %>%
  group_by(PlotID, scenario) %>%
  summarise(n_analog = n(),
            ave_elev_diff = mean(elev_diff),
            ave_km = mean(km))
biome.sup.plots.stats

write.csv(biome.sup.plots, file = "../output/biome_table.csv", row.names = FALSE)
write.csv(biome.sup.plots.stats, file = "../output/biome_analog_stats.csv", row.names = FALSE)
```


### FIA dataset

Load results files.  

```{r fia_results}

current.f.files <- list.files(path="../output/current",full.names=TRUE, pattern = "_FIA.txt")
current.f <- list()
for (file in current.f.files){
  name.tmp <- gsub(pattern = "../output/current/current_analog_", replacement = "", x = file)
  name <- gsub(pattern = "_FIA.txt", replacement = "", x = name.tmp)
  current.f[[name]] <- read.delim(file = file, header=T, sep="", na.strings="NA") %>%
    mutate(scenario = "current",
           location = name) 
}

future.f.files <- list.files(path="../output/future",full.names=TRUE, pattern = "_FIA.txt")
future.f <- list()
for (file in future.f.files){
  name.tmp <- gsub(pattern = "../output/future/future_analog_", replacement = "", x = file)
  name <- gsub(pattern = "_FIA.txt", replacement = "", x = name.tmp)
  future.f[[name]] <- read.delim(file = file, header=T, sep="", na.strings="NA") %>%
    mutate(scenario = "future",
           location = name) 
}

```

Count analogs.  

```{r fia2}
## Current analogs
# remove empty elements from list
current.f <- current.f[map(current.f, ~dim(.)[1]) > 0]
# bind rows from list to dataframe    
current.f.df <-  bind_rows(current.f, .id = "PlotID") %>%
  filter(Edist <= 0.84)
## Future analogs
# remove empty elements from list
future.f <- future.f[map(future.f, ~dim(.)[1]) > 0]
# bind rows from list to dataframe    
future.f.df <-  bind_rows(future.f, .id = "PlotID") %>%
  filter(Edist <= 0.84)

f.plot <- current.f.df %>%
  bind_rows(future.f.df) %>%
  select(PlotID, scenario, Edist, Edist_thresh, elev, elev_diff, km) 

f.bar <- ggplot(f.plot, aes(x=scenario, fill=Edist_thresh)) +
  geom_bar(position="dodge") +
  labs(title="FIA analog count") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) 
f.bar

```

Compare current and future analogs.

```{r fia3}


fia.sup.plots <- current.f.df %>%
  bind_rows(future.f.df) %>%
  distinct() %>%
 # filter(PlotID %in% plots) %>%
  select(-PointID, -long, -lat, -Edist, -Edist_thresh, -elev, -elev_diff, -km, -PointID_fia) %>%
  replace(is.na(.), 0) %>%
  group_by(scenario) %>% # if plots > 1, group by PlotID
  summarise(n = n(),
            across(where(is.numeric), sum)) %>%
  mutate(across(where(is.numeric) & !n, ~.x/n)) %>%
  pivot_longer(cols = where(is.numeric) & !n, names_to = "species_code", values_to = "freq") %>%
  select(-n) %>%
  pivot_wider(names_from = c("scenario"), values_from = freq) %>%
  mutate(remove = case_when(if_all(starts_with("current"), ~.x < .05) & if_all(starts_with("future"), ~.x < .05) ~ "drop",
                           TRUE ~ "keep")) %>%
  filter(remove == "keep") %>%
  select(-remove) %>%
  # arrange(desc(current), .by_group = TRUE)  %>%
#  replace(is.na(.), 0) %>% 
  left_join(fia.species.lookup)
fia.sup.plots


fia.sup.plots.stats <- current.f.df %>%
  bind_rows(future.f.df) %>%
  distinct() %>%
#  filter(PlotID %in% plots) %>%
  select(scenario, elev, elev_diff, km, PlotID) %>%
  group_by(PlotID, scenario) %>%
  summarise(n_analog = n(),
            ave_elev_diff = mean(elev_diff),
            ave_km = mean(km))
fia.sup.plots.stats

write.csv(fia.sup.plots, file = "../output/FIA_species_table.csv", row.names = FALSE)
write.csv(fia.sup.plots.stats, file = "../output/FIA_analog_stats.csv", row.names = FALSE)


```



## Session Info

```{r end}
sessionInfo()
```
